---
title: "Post-Processing QC"
author: "Dominik Schneider"
date: "`r strftime(Sys.Date(),'%b %d, %Y')`"
output:
  bookdown::html_document2: 
    df_print: paged
    fig_caption: yes
    fig_height: 16
    fig_width: 16
    toc: yes
    toc_depth: 3
    toc_float: yes
    self_contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, message=F, warning = F, fig.width=16)
library(tidyverse)
library(tidylog)
library(here)
library(cowplot)
library(scico)
```

```{r setup_plotdir}
datadir = here('output','from_diy_data')
figdir = here(datadir,'figs')
dir.create(figdir, rec=T)
```

# Read data from image analysis

```{r readdata, include=T}
data = read_csv(file.path(datadir,'output_psII_level0.csv'), #reading the data from the image processing where we explicitly define what NA values look like
                na=c('nan','--') ) %>% 
  mutate(date = as.Date(date), #formatting dates properly makes plotting easier
         idate = date-min(date)+1, #days after treatment begins
         treatment = as_factor(treatment), #creating factors will make plotting order easier to change
         gtype = toupper(gtype), #unifying all mutant codes uppercase so that wildtype is  'WT'
         gtype = as_factor(gtype),
         gtype = fct_relevel(gtype,sort),
         gtype = fct_relevel(toupper(gtype),'WT',after=0), #putting WT to the front of the list so we can color it black always
         measurement = case_when(parameter == 'FvFm' ~ 'FvFm', #this will help group the different PSII parameters between Fv/Fm and Induction Curve
                                 TRUE ~ 'IndC'),
         parameter = as_factor(parameter),
         plantarea = plantarea * 3.0625) # we should reprocess the images. original resolution guess was 0.2 mm/pixel. Measured res was 0.35. This is 3.0625 times more area per square pixel. remove this if psII.py is rerun with the right pixel resolution.
```

```{r gtypeColors, include=F}
# set your genotype color palette. 'black' is always WT 
gtypeColors = c('black',RColorBrewer::brewer.pal(length(levels(data$gtype))-1,'Set2'))
names(gtypeColors) <- levels(data$gtype)

italic_labels <- function(names){
  if(is.factor(names)){
    names = as.character(names)
  }
sapply(strsplit(names,'[. ]'), 
  function(x) {
    if( all(x != 'WT') & length(x)>1){
      x=tolower(x)
        # parse(text = paste0(x[1],".","italic(", tolower(x[2]), ")"))
        bquote(paste(.(x[1]),".",italic(.(x[2]))))
      # substitute(paste(treatment,".",italic(gtype), list(treatment = tolower(x[1]) , gtype=tolower(x[2]) ) ))
    } else if(all(x!='WT') & length(x)==1){
      x=tolower(x)
      substitute(italic(gtype), list(gtype=tolower(x) ) )
        # parse(text = paste0("italic(", tolower(x), ")"))
    } else if(length(x)>1){
        paste(x[1],x[2],sep='.')
    } else {
        x
    }
  })}

```

# Check image results

During image processing the detected objects are checked for:

  1. whether the plant is completely in frame (top row) 
  
  2. whether the plants are unique (bottom row)


```{r plot, echo=F, message=F, fig.height=8, fig.width=20}
data %>% 
  filter(parameter == 'FvFm') %>% 
  gather(qc, qcvalue, obj_in_frame:unique_roi) %>%
ggplot()+
  geom_tile(aes(x=sampleid,y=roi,fill=qcvalue),colour='grey80')+
  facet_grid(qc~date)+
  scale_y_continuous(limits=c(-1,9), breaks = seq(0,8), expand = c(0,0))+
  coord_fixed()+
  theme_cowplot(font_size = 18)+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))+
  labs(title = 'Image processing quality control checks',
       subtitle = 'FALSE means the quality check was tripped.\nTRUE means the quality check was not tripped.\nNA means there was no plant in the image.')


```

# Subset Valid data

For subsequent plots we will remove invalid datapoints based on the quality checks above. These are saved as the "level1" dataset.

```{r}
data <- 
  data %>% 
  filter(unique_roi == T, obj_in_frame == T) # we only want the observations that passed our tests.

write_csv(data, here('output/from_diy_data/output_psII_level1.csv'))

```


# Plant Area

```{r}
plantareadF = data %>% 
  filter(measurement == 'FvFm') %>% 
  group_by(gtype, treatment, idate,date, parameter, measurement) %>% 
  summarise(avg = mean(plantarea),
            stdev = sd(plantarea),
            sterr = stdev/sqrt(n()))

plantarea_ts =
  plantareadF %>% 
  ggplot(aes(x=idate))+
  geom_ribbon(aes(ymin=avg-sterr, ymax=avg+sterr, group=interaction(gtype,treatment)), fill='grey80')+
  geom_path(aes(x=idate, y=avg, color=gtype, linetype = treatment),size=1.2)+
  scale_color_manual(name = 'Genotype', values = gtypeColors, labels = italic_labels(names(gtypeColors)))+
  scale_linetype_discrete(name = 'Treatment')+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
  labs(title = 'Growth Curves',
       subtitle = 'Shaded region represents +/- 1 standard error',
       y = bquote('plant area [' * mm^2 *']'))+
  facet_wrap(~.)+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1),
        legend.justification = c(0,1),
        legend.position = c(0.05,0.96))

# plantarea_ts
```

```{r}
plantarea_errors <- 
  plantareadF %>% 
  ggplot()+
  geom_col(aes(x=idate, y=avg, fill=gtype), position=position_dodge2())+
  geom_errorbar(aes(x=idate, ymin = avg-sterr, ymax=avg+sterr), width=0.9, position=position_dodge2(), color='grey50')+
  facet_wrap(~treatment)+
  scale_fill_manual(values = gtypeColors, guide=F)+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
  # scale_y_continuous(limits = c(0,250))+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1),
        axis.text.y = element_blank())+
  labs(title = 'Plant Area',
       subtitle = 'Error bars represent +/- 1 standard error',
       y = '')
```

```{r, fig.height=8}
plot_grid(plantarea_ts,plantarea_errors)
```



# YII

       
```{r, fig.width=18, fig.height=8}
yiidF = data %>% 
  filter(frame %in% c('Fm','Fmp')) %>% 
  group_by(gtype, treatment, idate, date, parameter, measurement) %>% 
  summarise(avg = mean(yii_avg),
            stdev = sd(yii_avg),
            sterr = stdev/sqrt(n()))

ggplot(yiidF %>% filter(parameter == 'FvFm')) +
  geom_col(aes(y = avg, x=gtype, fill=treatment), position=position_dodge2())+
  geom_errorbar(aes(x=gtype, ymin=avg-sterr, ymax=avg+sterr), position=position_dodge2(), color='grey50')+
  scale_fill_grey()+
  facet_grid(~date)+
  labs(title = 'Fv/Fm',
       subtitle = 'Error bars represent +/- 1 standard error')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.4, hjust=1))
```


```{r, fig.height=8}
fvfm_ts = ggplot(yiidF %>% filter(parameter == 'FvFm'), aes(x=idate)) +
  geom_ribbon(aes(ymin=avg-sterr, ymax=avg+sterr, group=interaction(gtype,treatment)), fill='grey80')+
  geom_line(aes(y = avg, color=gtype, linetype=treatment))+
  labs(title = 'Fv/Fm',
       subtitle = 'Shaded region represents +/- 1 standard error',
       y = 'average')+
  scale_color_manual(values=gtypeColors)+
    # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+ 
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))

fvfm_ts
```

```{r, fig.height=10}
ggplot() + 
  geom_line(data = yiidF %>% filter(parameter != 'FvFm'), aes(x = parameter, y = avg, group = interaction(treatment, gtype), linetype = treatment, color = gtype), size=1)+
  facet_wrap(~date)+
  scale_color_manual(values = gtypeColors)+
  scale_x_discrete(drop = TRUE)+
  scale_linetype_discrete(limits=c('control','fluc'))+
  scale_shape_discrete(limits = c('control','fluc'))+
  labs(title = 'YII during Induction Curves')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.4, hjust=1))
```

## Steady-state YII

```{r, fig.height=8}

yii_ss = data %>% 
  filter(frame %in% c('Fm','Fmp'), 
         parameter %in% levels(parameter)[(nlevels(parameter)-1):nlevels(parameter)]) %>%
  group_by(gtype, treatment, date, idate) %>% 
  summarise(avg = mean(yii_avg),#average last two values from Induction Curve
            stdev = sd(yii_avg),
            sterr = stdev/sqrt(n()))

ggplot(yii_ss) +
  geom_col(aes(y = avg, x=gtype, fill=treatment), position=position_dodge2())+
  geom_errorbar(aes(x=gtype, ymin=avg-sterr, ymax=avg+sterr),position=position_dodge2(), color='grey50', width=0.7)+
  scale_fill_grey()+
  facet_grid(~date)+
  labs(title = 'YII (end of Induction Curve)',
       subtitle = 'Statistic are averaged from the last two data points on the Induction Curve\nError bars represent +/- 1 standard error',
       y = 'average')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1))

```

```{r, fig.height=8}
yii_ss_ts <- 
  ggplot(yii_ss, aes(x=idate)) +
  geom_ribbon(aes(ymin=avg-sterr, ymax=avg+sterr, group=interaction(gtype,treatment)), fill='grey80')+
  geom_line(aes(y = avg, color=gtype, linetype=treatment))+
  labs(title = 'Steady-state YII (end of Induction Curve)',
       subtitle = 'Average of last two data points on the Induction Curve\nError bars represent +/- 1 standard error',
       y = 'average')+
  scale_color_manual(values=gtypeColors)+
    # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))

yii_ss_ts
```

# NPQ

```{r, fig.height=10}
npqdF = data %>% 
  filter(frame %in% c('Fm','Fmp')) %>% 
  group_by(gtype, treatment, idate, date, parameter, measurement) %>% 
  summarise(avg = mean(npq_avg),
            stdev = sd(npq_avg))


ggplot() + 
  geom_line(data = npqdF %>% filter(parameter != 'FvFm'), aes(x = parameter, y = avg, group = interaction(treatment,gtype), linetype = treatment, color = gtype), size=1)+
  facet_wrap(~date)+
    scale_color_manual(values = gtypeColors)+
  scale_x_discrete(drop = TRUE)+
  scale_linetype_discrete(limits=c('control','fluc'))+
  scale_shape_discrete(limits = c('control','fluc'))+
  labs(title = 'NPQ during Induction Curves')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1))
```

## Steady-state NPQ

```{r, fig.height=8}
npq_ss = data %>% 
  filter(frame %in% c('Fm','Fmp'), parameter %in% levels(parameter)[(nlevels(parameter)-1):nlevels(parameter)]) %>% 
  group_by(gtype, treatment, idate, date) %>% 
  summarise(avg = mean(npq_avg),
            stdev = sd(npq_avg),
            sterr = stdev/sqrt(n()))

ggplot(npq_ss) +
  geom_col(aes(y = avg, x=gtype, fill=treatment), position=position_dodge2())+
  geom_errorbar(aes(x=gtype, ymin=avg-sterr, ymax=avg+sterr),position=position_dodge2(), color='grey50', width=0.7)+
  scale_fill_grey()+
  facet_grid(~date)+
  labs(title = 'NPQ (end of Induction Curve)',
       subtitle = 'Statistic are averaged from the last two data points on the Induction Curve\nError bars represent +/- 1 standard error',
       y = 'average')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1))

```
```{r, fig.height=8}
npq_ss_ts <- 
  ggplot(npq_ss, aes(x=idate)) +
  geom_ribbon(aes(ymin=avg-sterr, ymax=avg+sterr, group=interaction(gtype,treatment)), fill='grey80')+
  geom_line(aes(y = avg, color=gtype, linetype=treatment))+
  labs(title = 'Steady-state NPQ (end of Induction Curve)',
       subtitle = 'Average of last two data points on the Induction Curve\nError bars represent +/- 1 standard error',
       y = 'average')+
  scale_color_manual(values=gtypeColors)+
    # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))

npq_ss_ts

```


# Treatment Effects

## How does each genotype respond to the treatment compared to control (steady light) conditions?

### Plant Area
```{r}
plantarea_controlavg <- 
  plantareadF %>% 
  ungroup %>% 
  select(idate, date,treatment, gtype, baseavg = avg, basesd=stdev)

chgdF = full_join(plantareadF %>% filter(treatment == 'fluc'), 
                  plantarea_controlavg %>% filter(treatment == 'control') %>% select(-treatment), 
                  by=c('idate','date','gtype')) %>% 
  mutate(chg = avg-baseavg)



```


```{r, fig.height=4, fig.width=8}
colorlimits_plantarea = max(abs(range(chgdF %>% pull(chg), na.rm=T)))

plantarea_raster_treatment <- 
  ggplot(chgdF, aes(x=idate))+
  geom_raster(aes(y=gtype,fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
    scale_y_discrete(limits = rev(levels(chgdF$gtype)),
                   labels = italic_labels(rev(levels(chgdF$gtype))),
                   expand=c(0,0), 
                   drop=T)+
  scale_fill_scico(name = 'Deviation\n(Modulus power 2)',
                   palette = 'tofino', 
                   limits = c(-colorlimits_plantarea, colorlimits_plantarea)*1.3, 
                   oob = scales::squish, 
                   trans = scales::modulus_trans(p=2))+ 
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Treatment effects on plant area compared to treatment control",
       subtitle = "Positive values indicate an increase under fluctuating light conditions")

plantarea_raster_treatment

```


### YII

```{r, include = F}
yii_controlavg = yiidF %>%
  ungroup %>% 
  select(idate,date,treatment, parameter, gtype, baseavg = avg, basesd=stdev)

chgdF = full_join(yiidF %>% filter(treatment == 'fluc'), 
                  yii_controlavg %>% filter(treatment == 'control')%>% select(-treatment), 
                  by=c('idate','date','parameter','gtype')) %>% 
  mutate(chg = avg-baseavg)

```

```{r}
fvfmchg = data %>% 
  select(sampleid, treatment, roi, idate, date, gtype, parameter, yii_avg) %>% 
  filter(parameter == 'FvFm') %>% 
  full_join(yii_controlavg %>% filter(treatment=='control', parameter == 'FvFm') %>% select(-treatment)) %>% 
  mutate(chg = yii_avg - baseavg)
         # chg = ifelse(abs(chg) < basesd,NA,chg))

colorlimits = max(abs(range(fvfmchg$chg, na.rm=T)))

segloc <- 
  fvfmchg %>%
  mutate(x = min(date),
         xend = max(date)) %>% 
  distinct(gtype, sampleid,roi, x,xend) %>% 
  arrange(gtype, sampleid, roi) %>% 
  mutate(lag = lag(gtype),
         loc = ifelse(lag!=gtype,row_number(),NA) ) %>% 
  drop_na() 


ggplot(fvfmchg %>% arrange(sampleid, roi), aes(x=idate))+
  geom_raster(aes(y= interaction(gtype, treatment, sampleid, roi, lex.order = T, sep=':'), fill=chg))+
  geom_hline(data = segloc, aes(yintercept =  loc-0.5))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
  # scale_fill_scico(palette = 'broc', limits=c(-colorlimits, colorlimits), oob = scales::squish, direction=1)+
  scale_fill_distiller(palette = 'PuOr',limits=c(-colorlimits, colorlimits), oob = scales::squish)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Fv/Fm: Treatment effect compared to treatment control",
       subtitle = 'mean control phenotype for each genotype from each day substracted from daily measurement',
       y = 'replication id')

```

```{r, fig.height=4, fig.width=8}

colorlimits_fvfm = max(abs(range(chgdF %>% filter(parameter=='FvFm') %>% pull(chg), na.rm=T)))

fvfm_raster_treatment <- 
  ggplot(chgdF %>% filter(parameter=='FvFm'), aes(x=idate))+
  geom_raster(aes(y=gtype,fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
  scale_y_discrete(limits = rev(levels(chgdF$gtype)),
                   labels = italic_labels(rev(levels(chgdF$gtype))),
                    expand = expand_scale(0,0))+
  scale_fill_scico(name = 'Deviation\n(Modulus base 10)', 
                   palette = 'lisbon', 
                   limits=c(-colorlimits_fvfm, colorlimits_fvfm)*1.5, 
                   oob = scales::squish,
                   trans = scales::modulus_trans(p=10))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Treatment effects on Fv/Fm compared to treatment control",
       subtitle = "Positive values indicate an increase under fluctuating light conditions")

fvfm_raster_treatment
```



```{r, fig.height = 12}
ggplot(chgdF %>% filter(parameter != 'FvFm'), aes(x = idate))+
  geom_line(aes( y = chg, color=gtype, group=gtype))+
  facet_wrap(~parameter)+
    scale_color_manual(values = gtypeColors)+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
labs(title = 'Treatment effects based on the Induction Curve',
       subtitle = 'Positive values indicate an increase under fluctuating light conditions')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```

```{r, fig.height = 10}
ggplot() + 
  geom_line(data = chgdF %>% filter(parameter != 'FvFm'), aes(x = parameter, y = chg, group = gtype,  color = gtype))+
  facet_wrap(~date)+
  scale_color_manual(values = gtypeColors)+
  scale_x_discrete(drop = TRUE)+
  scale_linetype_discrete(limits=c('control','fluc'))+
  scale_shape_discrete(limits = c('control','fluc'))+
  labs(title = 'Treatment effects based on the Induction Curve',
       subtitle = 'Positive values indicate an increase under fluctuating light conditions')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```

#### Steady-state YII


```{r, fig.height=4, fig.width=8}

controlavg = yii_ss %>%
  ungroup %>% 
  filter(treatment=='control') %>% 
  select(idate, date, gtype, baseavg = avg, basesd = stdev)

chgdF = full_join(yii_ss %>% filter(treatment == 'fluc'), 
                  controlavg) %>% 
        mutate(chg = avg-baseavg)


colorlimits = max(abs(range(chgdF$chg, na.rm=T)))

yii_ss_raster_treatment <- 
  ggplot(chgdF, aes(x=idate))+
  geom_raster(aes(y=gtype,fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
    scale_y_discrete(limits = rev(levels(chgdF$gtype)),
                   labels = italic_labels(rev(levels(chgdF$gtype))),
                      expand=c(0,0))+
    scale_fill_scico(name = 'Deviation\n(Modulus base 10)', 
                   palette = 'lisbon', 
                   limits=c(-colorlimits, colorlimits)*1.2, 
                   oob = scales::squish,
                   trans = scales::modulus_trans(p=10))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Treatment effects on steady-state YII compared treatment control",
       subtitle = "Positive values indicate an increase under fluctuating light conditions.")

yii_ss_raster_treatment

colorlimits_ssyii <- colorlimits

```


### NPQ

```{r, fig.width = 16}
npq_controlavg = npqdF %>%
  ungroup %>% 
  select(idate, date,treatment, parameter, gtype, baseavg = avg, basesd=stdev)

chgdF = full_join(npqdF %>% filter(treatment == 'fluc'), 
                  npq_controlavg %>% filter(treatment == 'control')%>% select(-treatment), 
                  by=c('idate','date','parameter','gtype')) %>% 
  mutate(chg = avg-baseavg) %>% 
  filter(parameter != 'FvFm')
```

```{r, fig.height=12}
ggplot(chgdF, aes(x=idate))+
  geom_line(aes(y = chg, color=gtype, group=gtype))+
  facet_wrap(~parameter)+
    scale_color_manual(values = gtypeColors)+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
labs(title = 'Treatment effects on NPQ based on Induction Curve',
       subtitle = 'Positive values indicate an increase under fluctuating light conditions')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.4, hjust=1))

```

```{r, fig.height=10}
ggplot() + 
  geom_line(data = chgdF , aes(x = parameter, y = chg, group = gtype,  color = gtype))+
  facet_wrap(~date)+
  scale_color_manual(values = gtypeColors)+
  scale_x_discrete(drop = TRUE)+
  scale_linetype_discrete(limits=c('control','fluc'))+
  scale_shape_discrete(limits = c('control','fluc'))+
  labs(title = 'Difference in NPQ between fluctuating and steady light',
       subtitle = 'Positive values indicates increased NPQ with fluctuating light')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.4, hjust=1))
```

#### Steady-state NPQ

```{r, fig.height=4, fig.width=8}

controlavg = npq_ss %>%
  ungroup %>% 
  filter(treatment == 'control') %>% 
  select(idate,date, gtype, baseavg = avg, basesd = stdev)

chgdF = full_join(npq_ss %>% filter(treatment == 'fluc'), 
                  controlavg) %>% 
  mutate(chg = avg-baseavg)


colorlimits = max(abs(range(chgdF$chg, na.rm=T)))

npq_ss_raster_treatment <- 
  ggplot(chgdF, aes(x=idate))+
  geom_raster(aes(y=gtype,fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
    scale_y_discrete(limits = rev(levels(chgdF$gtype)),
                   labels = italic_labels(rev(levels(chgdF$gtype))),
                                      expand=c(0,0))+
    scale_fill_scico(name = 'Deviation\n(Modulus power 6)', 
                   palette = 'berlin', 
                   limits=c(-colorlimits, colorlimits), 
                   oob = scales::squish,
                   trans = scales::modulus_trans(p=6))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Treatment effects on steady-state NPQ compared to treatment control",
       subtitle = "Positive values indicate an increase under fluctuating light conditions.")

npq_ss_raster_treatment

colorlimits_ssnpq <- colorlimits
```

## How does each mutant respond to the treatment compared to WT?

### Plant Area
```{r}
plantarea_WTavg = plantareadF %>%
  ungroup %>% 
  filter(gtype=='WT') %>% 
  select(idate, date, treatment, -gtype, baseavg = avg, basesd=stdev)

chgdF = full_join(plantareadF, 
                  plantarea_WTavg, 
                  by=c('idate','date','treatment')) %>% 
  mutate(chg = avg-baseavg) %>% 
  filter(gtype!='WT')


```

```{r, fig.height=4, fig.width=8}

colorlimits_plantarea = max(colorlimits_plantarea, max(abs(range(chgdF$chg, na.rm=T))))

plantarea_raster_wt <- 
  ggplot(chgdF, aes(x=idate))+
  geom_raster(aes(y=interaction(treatment,gtype),fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
  scale_y_discrete(limits = rev(levels(fct_drop(interaction(chgdF$treatment, chgdF$gtype)))),
                   labels = italic_labels(rev(levels(fct_drop(interaction(chgdF$treatment, chgdF$gtype))))),
                   expand=c(0,0),
                   drop=T)+
    scale_fill_scico(name = 'Deviation\n(Modulus power 2)', 
                   palette = 'tofino', 
                   limits=c(-colorlimits_plantarea, colorlimits_plantarea)*1.3, 
                   oob = scales::squish,
                   trans = scales::modulus_trans(p=2))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
 labs(title="Treatment effects on plant area compared to WT",
       subtitle = "Positive values indicate an increase compared to WT")

plantarea_raster_wt
```



### YII

```{r}
yii_WTavg = yiidF %>%
  ungroup %>% 
  filter(gtype=='WT') %>% 
  select(idate, date, treatment, parameter, -gtype, baseavg = avg, basesd=stdev)

chgdF = full_join(yiidF , 
                  yii_WTavg, 
                  by=c('idate','date','parameter','treatment')) %>% 
  mutate(chg = avg-baseavg) %>% 
  filter(gtype!='WT')
```

```{r}
fvfmchg = data %>% 
  select(sampleid, treatment, roi, idate, date, gtype, parameter, yii_avg) %>% 
  filter(parameter == 'FvFm') %>% 
  full_join(yii_WTavg %>% filter(parameter == 'FvFm')) %>% 
  mutate(chg = yii_avg - baseavg)
         # chg = ifelse(abs(chg) < basesd,NA,chg))

colorlimits = max(abs(range(fvfmchg$chg, na.rm=T)))

segloc <- 
  fvfmchg %>%
  mutate(x = min(date),
         xend = max(date)) %>% 
  distinct(gtype, sampleid,roi, x,xend) %>% 
  arrange(gtype, sampleid, roi) %>% 
  mutate(lag = lag(gtype),
         loc = ifelse(lag!=gtype,row_number(),NA) ) %>% 
  drop_na() 


ggplot(fvfmchg %>% arrange(sampleid, roi), aes(x=idate))+
  geom_raster(aes(y= interaction(gtype, treatment, sampleid, roi, lex.order = T, sep=':'), fill=chg))+
  geom_hline(data = segloc, aes(yintercept =  loc-0.5))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
  # scale_y_discrete(limits=rev(interaction(fvfmchg$gtype, fvfmchg$treatment, fvfmchg$sampleid, fvfmchg$roi, lex.order = T, sep=':')))+ #this doesn't work
  # scale_fill_scico(palette = 'broc', limits=c(-colorlimits, colorlimits), oob = scales::squish, direction=1)+
  scale_fill_distiller(palette = 'PuOr',limits=c(-colorlimits, colorlimits), oob = scales::squish)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Fv/Fm: Treatment effect compared to WT",
       subtitle = 'mean WT phenotype for each treatment from each day substracted from daily measurement',
       y = 'replication id')

```

```{r, fig.height=4, fig.width=8}
colorlimits_fvfm = max(colorlimits_fvfm, max(abs(range(chgdF %>% filter(parameter=='FvFm') %>% pull(chg), na.rm=T))))

fvfm_raster_wt <- 
  ggplot(chgdF %>% filter(parameter=='FvFm'), aes(x=idate))+
  geom_raster(aes(y=interaction(treatment, gtype),fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
  scale_y_discrete(limits = rev(levels(fct_drop(interaction(chgdF$treatment, chgdF$gtype)))),
                   labels = italic_labels(rev(levels(fct_drop(interaction(chgdF$treatment, chgdF$gtype))))),
                   expand=c(0,0))+
scale_fill_scico(name = 'Deviation\n(Modulus power 10)', 
                   palette = 'lisbon', 
                   limits=c(-colorlimits_fvfm, colorlimits_fvfm)*1.5, 
                   oob = scales::squish,
                   trans = scales::modulus_trans(p=10))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Treatment effects on Fv/Fm compared to WT",
       subtitle = "Positive values indicate an increase compared to WT")

fvfm_raster_wt
```

```{r, fig.height = 12 }
ggplot(chgdF %>% filter(parameter != 'FvFm'), aes(x=idate))+
  geom_line(aes(y = chg, color=gtype, linetype = treatment))+
  facet_wrap(~parameter)+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
  scale_color_manual(values = gtypeColors)+
  labs(title = 'Treatment effects based on the Induction Curve',
       subtitle = 'Positive values indicate an increase compared to WT')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```

```{r, fig.height = 10}
ggplot() + 
  geom_line(data = chgdF %>% filter(parameter != 'FvFm'), aes(x = parameter, y = chg, linetype=treatment,  color = gtype, group=interaction(gtype,treatment)))+
  facet_wrap(~date)+
  scale_color_manual(values = gtypeColors)+
  scale_x_discrete(drop = TRUE)+
  scale_linetype_discrete(limits=c('control','fluc'))+
  scale_shape_discrete(limits = c('control','fluc'))+
  labs(title = 'Treatment effects based on the Induction Curve',
       subtitle = 'Positive values indicate an increase compared to WT')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```

#### Steady-state YII


```{r, fig.height=4, fig.width=8}

WTavg = yii_ss %>%
  ungroup %>% 
  filter(gtype=='WT') %>% 
  select(date, treatment, baseavg = avg, basesd = stdev)

chgdF = full_join(yii_ss, 
                  WTavg, by=c('date','treatment')) %>% 
  mutate(chg = avg-baseavg) %>% 
  filter(gtype!='WT')


colorlimits = max(abs(range(chgdF$chg, na.rm=T)))
yii_ss_raster_wt <- 
  ggplot(chgdF, aes(x=idate))+
  geom_raster(aes(y=interaction(treatment,gtype),fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
  scale_y_discrete(limits = rev(levels(fct_drop(interaction(chgdF$treatment, chgdF$gtype)))),
                   labels = italic_labels(rev(levels(fct_drop(interaction(chgdF$treatment, chgdF$gtype))))),
                    expand=c(0,0))+
scale_fill_scico(name = 'Deviation\n(Modulus power 10)', 
                   palette = 'lisbon', 
                   limits=c(-colorlimits, colorlimits)*1.2,
                   oob = scales::squish,
                   trans = scales::modulus_trans(p=10))+  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Treatment effects on steady-state YII compared to WT",
       subtitle = 'Positive values indicate an increase compared to WT')

yii_ss_raster_wt
colorlimits_ssyii = max(colorlimits, colorlimits_ssyii)
```


### NPQ

```{r}
npq_WTavg = npqdF %>%
  ungroup %>% 
  filter(gtype == 'WT') %>% 
  select(idate, date, treatment, parameter, -gtype, baseavg = avg, basesd = stdev)

chgdF = full_join(npqdF, 
                  npq_WTavg, 
                  by=c('idate','date','parameter','treatment')) %>% 
  mutate(chg = avg-baseavg) %>% 
  filter(parameter != 'FvFm')
```

```{r, fig.height = 12}
ggplot(chgdF, aes(x=idate))+
  geom_line(aes(y = chg, color = gtype, linetype = treatment))+
  facet_wrap(~parameter)+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d')+
  scale_color_manual(values = gtypeColors)+
  labs(title = 'Treatment effects based on the Induction Curve',
       subtitle = 'Positive values indicate an increase compared to WT')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```

```{r, fig.height = 10}
ggplot() + 
  geom_line(data = chgdF, aes(x = parameter, y = chg, linetype=treatment,  color = gtype, group=interaction(gtype,treatment)))+
  facet_wrap(~date)+
  scale_color_manual(values = gtypeColors)+
  scale_x_discrete(drop = TRUE)+
  scale_linetype_discrete(limits=c('control','fluc'))+
  scale_shape_discrete(limits = c('control','fluc'))+
  labs(title = 'Treatment effects based on the Induction Curve',
       subtitle = 'Positive values indicate an increase compared to WT')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```

#### Steady-state NPQ

```{r, fig.height=4, fig.width=8}
WTavg = npq_ss %>%
  ungroup %>% 
  filter(gtype=='WT') %>% 
  select(idate, date, treatment, baseavg = avg, basesd = stdev)

chgdF = full_join(npq_ss, 
                  WTavg, 
                  by=c('idate','date','treatment')) %>% 
  mutate(chg = avg-baseavg) %>% 
  filter(gtype!='WT')


colorlimits = max(abs(range(chgdF$chg, na.rm=T)))

npq_ss_raster_wt <- 
  ggplot(chgdF, aes(x=idate))+
  geom_raster(aes(y=interaction(treatment,gtype),fill=chg))+
  # scale_x_date(date_breaks = '1 day', date_labels = '%b-%d', expand=c(0,0))+
  scale_y_discrete(drop=T,
                   limits=rev(levels(fct_drop(interaction(chgdF$treatment,chgdF$gtype)))),
                   labels=italic_labels(rev(levels(fct_drop(interaction(chgdF$treatment,chgdF$gtype))))),
                   expand=c(0,0))+
scale_fill_scico(name = 'Deviation\n(Modulus power 6)', 
                   palette = 'berlin', 
                   limits=c(-colorlimits, colorlimits)*1.2, 
                   oob = scales::squish,
                   trans = scales::modulus_trans(p=6))+  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Treatment effects on steady-state NPQ compared to WT",
       subtitle = 'Positive values indicate an increase compared to WT')

npq_ss_raster_wt

colorlimits_ssnpq = max(colorlimits_ssnpq, colorlimits)
```



# Compilation for Manuscript

```{r, include=T, fig.width=20, fig.height=15}

col1 = plot_grid(
  plantarea_ts+
     facet_null()+
    theme_cowplot(font_size=10)+
     theme(legend.just = c(0,1),
           legend.position = c(0.05,0.95),
           legend.spacing = unit(0,'npc'),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5),
           plot.title = element_text(face='plain', size=14))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0.5)),
  fvfm_ts +
      guides(color=FALSE, linetype=F) + 
        theme_cowplot(font_size=10)+
      theme(axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5),
            plot.title = element_text(face='plain', size=14))+
      scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0.5)),
  yii_ss_ts + 
    guides(color=FALSE, linetype=F) + 
        theme_cowplot(font_size=10)+
    theme(axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5),
                 plot.title = element_text(face='plain', size=14))+
    scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0.5)),
  npq_ss_ts +
    guides(color=FALSE, linetype=F) + 
        theme_cowplot(font_size=10)+
    theme(axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5),
         plot.title = element_text(face='plain', size=14))+
    scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0.5)),
  ncol=1, 
  align = 'hv',
  labels = LETTERS[c(2,5,8,11)],
  label_size = 30,
    label_fontface = 'plain'
)

col2 = plot_grid(
  plantarea_raster_treatment+
     guides(fill=F)+
     theme(axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_fill_scico(name = 'Deviation\n(Modulus power 2)', 
           palette = 'tofino', 
           limits=c(-colorlimits_plantarea, colorlimits_plantarea)*1.3, 
           oob = scales::squish,
           trans = scales::modulus_trans(p=2))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
   fvfm_raster_treatment+
     guides(fill=F)+
     scale_fill_scico(name = 'Deviation\n(Modulus base 10)', 
                      palette = 'lisbon', 
                      limits=c(-colorlimits_fvfm, colorlimits_fvfm)*1.5, 
                      oob = scales::squish,
                      trans = scales::modulus_trans(p=10))+
     theme(axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
   yii_ss_raster_treatment+guides(fill=F)+
     theme(axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_fill_scico(name = 'Deviation\n(Modulus base 10)', 
         palette = 'lisbon', 
         limits=c(-colorlimits_ssyii, colorlimits_ssyii)*1.2, 
         oob = scales::squish,
         trans = scales::modulus_trans(p=10))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
   npq_ss_raster_treatment+guides(fill=F)+
     theme(axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_fill_scico(name = 'Deviation\n(Modulus power 6)', 
         palette = 'berlin', 
         limits=c(-colorlimits_ssnpq, colorlimits_ssnpq)*1.1, 
         oob = scales::squish,
         trans = scales::modulus_trans(p=6))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
  ncol=1, 
  align = 'hv',
  labels = LETTERS[c(3,6,9,12)],
  label_size = 30,
  label_fontface = 'plain',
  hjust = 0)

col3 = plot_grid(
  plantarea_raster_wt+
     theme(legend.position = 'right', 
           legend.direction = 'vertical',
           legend.box = 'vertical',
           legend.box.just = 'left',
           legend.spacing = unit(0,'npc'),
           axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_fill_scico(name = 'Deviation\n(Modulus power 2)', 
           palette = 'tofino', 
           limits=c(-colorlimits_plantarea, colorlimits_plantarea)*1.3, 
           oob = scales::squish,
           trans = scales::modulus_trans(p=2))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
   fvfm_raster_wt+
     theme(legend.position = 'right', 
           legend.direction = 'vertical',
           legend.box = 'vertical',
           legend.box.just = 'left',
           legend.spacing = unit(0,'npc'),
           axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_fill_scico(name = 'Deviation\n(Modulus base 10)', 
           palette = 'lisbon', 
           limits=c(-colorlimits_fvfm, colorlimits_fvfm)*1.5, 
           oob = scales::squish,
           trans = scales::modulus_trans(p=10))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
   yii_ss_raster_wt+
     theme(legend.position = 'right', 
           legend.direction = 'vertical',
           legend.box = 'vertical',
           legend.box.just = 'left',
           legend.spacing = unit(0,'npc'),
           axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_fill_scico(name = 'Deviation\n(Modulus base 10)', 
           palette = 'lisbon', 
           limits=c(-colorlimits_ssyii, colorlimits_ssyii)*1.2, 
           oob = scales::squish,
           trans = scales::modulus_trans(p=10))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
   npq_ss_raster_wt+
     theme(legend.position = 'right', 
           legend.direction = 'vertical',
           legend.box = 'vertical',
           legend.box.just = 'left',
           legend.spacing = unit(0,'npc'),
           axis.title.y = element_blank(),
           axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5))+
     scale_fill_scico(name = 'Deviation\n(Modulus power 6)', 
           palette = 'berlin', 
           limits=c(-colorlimits_ssnpq, colorlimits_ssnpq)*1.1, 
           oob = scales::squish,
           trans = scales::modulus_trans(p=6))+
     scale_x_continuous(name = 'Day of Treatment', breaks = unique(data$idate), expand=c(0,0)),
    ncol=1,
    align = 'hv',
    labels = LETTERS[c(4,7,10,13)],
    label_size = 30,
    label_fontface = 'plain')


compfig = plot_grid(col1,col2,col3, rel_widths = c(1,1,1.25), ncol=3, axis='b')
compfig
cowplot::save_plot(figdir,'manuscript_fig7.pdf'),compfig, nrow=4, ncol=3, base_width=7, limitsize=F)
```



